Quasi-Monte Carlo Methods: What, Why, and How?

### 2025-06-24 response_letter.tex

The authors agree with all minor comments  flagged by the reviewer and thus have been rectified.

\newpage

\section*{Response to Reviewer 2}

\begin{response}{1.}
    Add a simple but practical application, like option pricing under Black-Scholes model, as an illustrative example. This reviewer finds the illustrative example in Section 2 lacks practical relevance
\end{response}
We thank the reviewer for the suggestion to include a more practical application. We have added an illustrative example on the fair price of an Asian option using MC and QMC methods to tie in with the discussions of variable transforms and control variates (see below).

@@

\begin{response}{4.}
    The last line on page 9 is just a reference...
\end{response}
With some rewriting, this is no longer an issue.

\begin{response}{5.}
    First line in Section 3.4: The acronym qMC is defined and used many times, why not use it here?
\end{response}
    Thank you. Acronym now used here.

\begin{response}{6.}
    Figure 9 on page 13. The $x_0 = \Delta$ equation in the figures overlaps with some markers, which makes it hard to see. Please reposition it.
\end{response}
Done.

\begin{response}{7.}
    The legends in Figure 12 and Figure 13 are too small
\end{response}
We have increased the font size of the legends for these figures.

### 2025-06-24 HickernellMCQMC2024_QMCTutorial_rev_1.tex
Updates from Overleaf

%%Fred's temporary macros
\usepackage[dvipsnames]{xcolor}
\newcommand{\FJHNote}[1]{{\textcolor{ForestGreen}{#1}}}
\newcommand{\NKNote}[1]{{\textcolor{blue}{ #1}}}
\newcommand{\AGSNote}[1]{{\textcolor{magenta}{#1}}}
%\usepackage[normalem]{ulem}

@@

\mainmatter              % start of a contribution
%
\title{Quasi-Monte Carlo Methods:  What, Why, and How?}
%
\titlerunning{Quasi-Monte Carlo Methods}  % abbreviated title (for running head)
%                                     also used for the TOC unless
%                                     \toctitle is used
%
\author{Fred J. Hickernell\inst{1,2} \and Nathan Kirk\inst{1,3} \and Aleksei G. Sorokin\inst{1,4}}
%
\authorrunning{Hickernell, Kirk, and Sorokin} % abbreviated author list (for running head)
%
%%%% list of authors for the TOC (use if author list has to be modified)
\tocauthor{Fred J. Hickernell, Nathan Kirk, Aleksei G. Sorokin}
%
\institute{
Department of Applied Mathematics, Illinois Institute of Technology, \\ Chicago, IL, 60616, USA\\
\and
\email{hickernell@iit.edu}, \
%\texttt{http://www.iit.edu/\textasciitilde{}hickernell}
\url{http://www.iit.edu/~hickernell}
\and
\email{nkirk@iit.edu}
\and
\email{asorokin@hawk.iit.edu}
}

@@

To understand how well these nodes are evenly distributed over $[0,1)^d$, imagine a box of the form
\begin{multline} \label{eq:element_box}
	[a_1 b^{-k_1}, (a_1 + 1) 2^{-k_1}) \times \cdots \times [a_d b^{-k_d}, (a_d + 1) 2^{-k_d}), \\
	a_j \in \{0, \ldots, 2^{k_j} - 1\}, \ k_j \in \{0,1, \ldots \}, \qquad j =1, \ldots d, \ \bsk \in \NN_0.
\end{multline}
This box has volume $2^{-(k_1 + \cdots +k_d)} = 2^{-\lVert\bsk\rVert_1}$.  For a digital net with $2^m$ nodes and $\lVert \bsk \rVert_1 \le m$, a ``fair share" of nodes for this box would be $2^{m- \lVert \bsk \rVert_1}$.  This will occur if and only if
\FJHNote{
\begin{equation} \label{eq:tcond}
\begin{aligned}
	\text{the first $k_1$ rows of the first $m$ columns of $\mathsf{C}_1$} & \text{ plus} \\
	\text{the first $k_2$ rows of the first $m$ columns of $\mathsf{C}_2$} & \text{ plus} \\
	 \vdots \hspace{10ex} & \\
	\text{the first $k_d$ rows of the first $m$ columns of $\mathsf{C}_d$} &\\
	\text{are linearly independent over addition and multiplication modulo $2$, }
\end{aligned}
\end{equation}}
regardless of the choice of $\bsa = (a_1, \ldots, a_d)$.  The $t$-value of a digital net with $2^m$ nodes is defined as \FJHNote{any} $t$ for which condition \eqref{eq:tcond} holds for all $\bsk$ with $\lVert \bsk \rVert_1 \le m-t$.  Equivalently, \FJHNote{for} this $t$, every box of the form \eqref{eq:element_box} with volume $2^{\lVert \bsk \rVert_1 - m}$ contains its fair share of $2^t$ nodes.  Such a digital net is then called a $(t,m,d)$-net.  An infinite sequence of the form \eqref{eq:digitalB} for which this condition holds for all non-negative $m$ is called a  $(t,d)$-sequence. \FJHNote{Ideally, one would like to construct $(t,m,d)$-nets and $(t,d)$-sequences with $t$ as small as possible.}

@@

When approximating the true $\mu$ with $\hat{\mu}_n$, stopping criterion determine how large $n$ should be so that the error $\lvert \mu - \hat{\mu}_n \rvert$ is below an error tolerance $\varepsilon$ either certainly or with sufficiently low uncertainty $\alpha$. The error tolerance and probability threshold are problem dependent. For example, in option pricing one may set the error tolerance to a penny, $\varepsilon=0.01$, and wish to find error bounds which hold with 95\% confidence, $\alpha=0.05$. The stopping criteria presented here are implemented in QMCPy \cite{QMCPy2020a}.

@@

One idea for constructing confidence intervals without replications is to approximate the decay rate of Fourier coefficients of $f$. Bounds of absolute certainty have be derived for cones of functions with reasonable decay of their Fourier coefficients \cite{HicEtal17a}. As in the qMC-CLT stopping criterion, one may iteratively double $n$ until the bounds are sufficiently tight.

@@

\begin{figure}
	\centering
 %    \includegraphics[width = 0.48\textwidth]{\figpath//NumSamples_Asian_Option_tol_0.001_d52.pdf}
	% \includegraphics[width = 0.48\textwidth]{\figpath/Times_Asian_Option_tol_0.001_d52.pdf} \quad 
    \includegraphics[width = 1\textwidth]{\figpath//Times_NumSamples_Asian_Option_tol_0.001_d52.pdf}
	\caption{\AGSNote{Comparison of Monte Carlo and Quasi-Monte Carlo stopping criteria for estimating the fair price of an Asian option with arithmetic mean. For QMC with Sobol' points, using a principal component analysis (PCA) decomposition of the Brownian motion provides substantial savings compared to the Cholesky decomposition. Using an Asian option with geometric mean control variate (CV) provides additional savings. \label{fig:asian_option_pda_vs_chol}}}
\end{figure}

@@

An example of this in quantitative finance is where one may use the payoff of a European option, whose expectation is known, as a control variate for more exotic options.  As noted in \cite{HicEtal03}, reducing the variation of $f$ via control variates may or may not resemble reducing the variance of $f$.  Control variates have been implemented in the stopping criteria based on Fourier coefficients described in the previous section \cite{HicEtal17a}. \AGSNote{In Figure \ref{fig:asian_option_pda_vs_chol} we show that, for pricing an Asian option with arithmetic mean, using an Asian option with geometric mean control variate can provide improved convergence.}

@@

>>>>

%%Fred's temporary macros
\usepackage[dvipsnames]{xcolor}
\newcommand{\FJHNote}[1]{{\textcolor{blue}{#1}}}
\newcommand{\NKNote}[1]{{\textcolor{blue}{ #1}}}
\newcommand{\AGSNote}[1]{{\textcolor{blue}{#1}}}
%\usepackage[normalem]{ulem}

@@

\mainmatter              % start of a contribution
%
\title{Quasi-Monte Carlo Methods:  What, Why, and How?}
%
\titlerunning{Quasi-Monte Carlo Methods}  % abbreviated title (for running head)
%                                     also used for the TOC unless
%                                     \toctitle is used
%
\author{Fred J. Hickernell\inst{1,2} \and Nathan Kirk\inst{1,3} \and Aleksei G. Sorokin\inst{1,4}}
%
\authorrunning{Hickernell, Kirk, and Sorokin} % abbreviated author list (for running head)
%
%%%% list of authors for the TOC (use if author list has to be modified)
\tocauthor{Fred J. Hickernell, Nathan Kirk, Aleksei G. Sorokin}
%
\institute{
Department of Applied Mathematics, Illinois Institute of Technology, \\ Chicago, IL, 60616, USA\\
\and
\email{hickernell@illinoistech.edu}, \
%\texttt{http://www.iit.edu/\textasciitilde{}hickernell}
\href{https://www.iit.edu/directory/people/fred-hickernell}{http://www.iit.edu/hickernell}
\and
\email{nkirk@illinoistech.edu}
\and
\email{asorokin@hawk.illinoistech.edu}
}

@@

To understand how well these nodes are evenly distributed over $[0,1)^d$, imagine a box of the form
\begin{multline} \label{eq:element_box}
	[a_1 b^{-k_1}, (a_1 + 1) 2^{-k_1}) \times \cdots \times [a_d b^{-k_d}, (a_d + 1) 2^{-k_d}), \\
	a_j \in \{0, \ldots, 2^{k_j} - 1\}, \ k_j \in \{0,1, \ldots \}, \qquad j =1, \ldots d, \ \bsk \in \NN_0.
\end{multline}
This box has volume $2^{-(k_1 + \cdots +k_d)} = 2^{-\lVert\bsk\rVert_1}$.  For a digital net with $2^m$ nodes and $\lVert \bsk \rVert_1 \le m$, a ``fair share" of nodes for this box would be $2^{m- \lVert \bsk \rVert_1}$.  This will occur if and only if
\FJHNote{
\begin{equation} \label{eq:tcond}
\begin{aligned}
	\text{the first $k_1$ rows of the first $m$ columns of $\mathsf{C}_1$} & \text{ plus} \\
	\text{the first $k_2$ rows of the first $m$ columns of $\mathsf{C}_2$} & \text{ plus} \\
	 \vdots \hspace{10ex} & \\
	\text{ the first $k_d$ rows of the first $m$ columns of $\mathsf{C}_d$} &\\
	\text{are linearly independent over addition and multiplication modulo $2$, }
\end{aligned}
\end{equation}}
regardless of the choice of $\bsa = (a_1, \ldots, a_d)$.  The $t$-value of a digital net with $2^m$ nodes is defined as \FJHNote{any} $t$ for which condition \eqref{eq:tcond} holds for all $\bsk$ with $\lVert \bsk \rVert_1 \le m-t$.  Equivalently, \FJHNote{for} this $t$, every box of the form \eqref{eq:element_box} with volume $2^{\lVert \bsk \rVert_1 - m}$ contains its fair share of $2^t$ nodes.  Such a digital net is then called a $(t,m,d)$-net.  An infinite sequence of the form \eqref{eq:digitalB} for which this condition holds for all non-negative $m$ is called a  $(t,d)$-sequence. \FJHNote{Ideally, one would like to construct $(t,m,d)$-nets and $(t,d)$-sequences with $t$ as small as possible.}

@@

When approximating the true $\mu$ with $\hat{\mu}_n$, stopping criterion determine how large $n$ should be so that the error $\lvert \mu - \hat{\mu}_n \rvert$ is below an error tolerance $\varepsilon$ either certainly or with sufficiently low uncertainty $\alpha$. The error tolerance and probability threshold are problem dependent. For example, in option pricing one may set the error tolerance to a penny, $\varepsilon=0.01$, and wish to find error bounds which hold with 95\% confidence, $\alpha=0.05$. The stopping criteria presented here are implemented in \texttt{qmcpy} \cite{QMCPy2020a}.

@@

One idea for constructing confidence intervals without replications is to approximate the decay rate of Fourier coefficients of $f$. Bounds of absolute certainty have be derived for cones of functions with reasonable decay of their Fourier coefficients \cite{HicEtal17a}. As in the qMC-CLT stopping criterion, one may iteratively double $n$ until the bounds are sufficiently tight. \AGSNote{Figure \ref{fig:asian_option_pda_vs_chol} compares these decay-tracking qMC algorithms against IID Monte Carlo stopping criterion across a variety of tolerances for pricing an Asian option from finance \cite[p.\ 8]{Gla03}.}

@@

\begin{figure}
	\centering
 %    \includegraphics[width = 0.48\textwidth]{\figpath//NumSamples_Asian_Option_tol_0.001_d52.pdf}
	% \includegraphics[width = 0.48\textwidth]{\figpath/Times_Asian_Option_tol_0.001_d52.pdf} \quad 
    \includegraphics[width = 1\textwidth]{\figpath//Times_NumSamples_Asian_Option_tol_0.001_d52.pdf}
	\caption{\AGSNote{Comparison of IID Monte Carlo and qMC stopping criteria for estimating the fair price of an Asian option with arithmetic mean. For QMC with Sobol' points, using a principal component analysis (PCA) decomposition of the Brownian motion provides substantial savings compared to the Cholesky decomposition. Using an Asian option with geometric mean control variate (CV) provides additional savings. \label{fig:asian_option_pda_vs_chol}}}
\end{figure}

@@

An example of this in quantitative finance is where one may use the payoff of a \AGSNote{similar} option, whose expectation is known, as a control variate for more exotic options.  As noted in \cite{HicEtal03}, reducing the variation of $f$ via control variates may or may not resemble reducing the variance of $f$.  Control variates have been implemented in the stopping criteria based on Fourier coefficients described in the previous section \cite{HicEtal17a}. \AGSNote{In Figure \ref{fig:asian_option_pda_vs_chol} we show that, for pricing an Asian option with arithmetic mean, using an Asian option with geometric mean control variate can provide improved convergence.}

@@

\subsubsection*{Acknowledgments.} This work is supported by National Science Foundation (DMS Grant No. 2316011). The third author would like to acknowledge support from the DOE Office of Science Graduate Student Research (SCGSR) Program.
